---
title: "070218 Research Work"
author: "Hanna Grossman"
date: "7/2/2018"
output: html_document
---
#finding PCA  - mukai's code  
```{r}
library(pls)
library(ggplot2)
library(dplyr)
expr.dat = readRDS("/Users/hannagrossman/Desktop/BDSI Research/expression.rds")
screen.dat = readRDS("/Users/hannagrossman/Desktop/BDSI Research/screening.rds")

pr.out <- prcomp(expr.dat)
names(pr.out)
pr.var = pr.out$sdev^2
pr.var[1:5]
pve = pr.var/sum(pr.var)
pve[1:5]
sum(pve[1:5])
pr_scale.out <- prcomp(expr.dat, scale=TRUE)
pr_scale.var = pr_scale.out$sdev^2
pr_scale.var[1:5]
pve_scale= pr_scale.var/sum(pr_scale.var)
pve_scale[1:5]
sum(pve_scale[1:10])
```

#choosing what drugs to use 
```{r}
head(screen.dat)
classes.1014 <- subset(screen.dat, DRUG_ID_lib == "1014")[,c("CL", "EFFECT")]
#246 observations 
head(classes.1014)
filter(classes.1014, EFFECT == TRUE) %>% nrow()
135/246

drug <- screen.dat$DRUG_ID_lib
unique(drug) 
#59 drugs 

#vector of all drug numbers 
a <- c(1001, 1003, 1004, 1006, 1010, 1011, 1012, 1016, 1017, 1018, 1020, 1022, 1023, 1024, 1025, 1028, 1031, 1032, 1033, 1042, 1053, 1057, 1058, 1059, 1149, 1007, 1014, 1039, 1049, 1061, 1062, 1066, 1067, 1052, 1129, 1038, 1054, 1060, 1009, 1043, 1037, 1013, 1026, 1029, 1030, 1036, 1047, 1050, 1069, 1008, 1072, 1015, 1019, 1021, 1046, 1170, 1133, 1175, 1005)

a[1]

for(i in a[1]:a[59]){
  x <- subset(screen.dat, DRUG_ID_lib == "i")["EFFECT"]
  print(nrow(x)) 
  print(filter(x, EFFECT == TRUE) %>% nrow())
}

screen.dat %>% group_by(DRUG_ID_lib, EFFECT) %>% summarise(count=n())
```


#seeing if classes.1014 match expr.1014 so we can then cbind them
```{r}
head(classes.1014)
expr.1014[1:6,1:6]

b <- rownames(expr.1014)
c <- classes.1014[,1]

d <- cbind(b,c)
d$CL <- as.numeric(d$CL)
d$b <- as.character(d$b)
d$b <- as.numeric(d$b)
d$e <- d$CL - d$b
```

#subsetting and cbinding
```{r}
classes.1014 <- subset(screen.dat, DRUG_ID_lib == "1014")[,c("CL","EFFECT")]
expr.1014 <- expr.dat[as.character(classes.1014$CL),]
head(classes.1014)
expr.1014[1:6,1:6]
exprnew <- cbind(classes.1014, expr.1014)
exprnew[1:6,1:6]
sum(is.na(exprnew))
dim(exprnew)
dim(expr.dat)
```

#PCA before kNN for loop below 
```{r}
#read in screen and expression data

#subset to drug 1014
classes.1014 <- subset(screen.dat, DRUG_ID_lib == "1014")[,c("CL", "EFFECT")]
expr.1014 <- expr.dat[as.character(classes.1014$CL),]
dim(expr.1014)
expr.1014[1:10,1:10]

#center the expression data by column means
center.expr <- scale(expr.1014)

#now do PCA - here we choose 3 components but we probably wann test 1-100 to see what is best - do loop 10 times (10 diff train and test sets)
pca <- svd(center.expr, nu=10, nv = 0)

pc.scores <- as.data.frame(diag(pca$d) %*% pca$u)
pc.scores <- cbind(classes.1014, pc.scores)
dim(pc.scores)
#head(pc.scores)
pc.scores[1:10,1:10]
```
#?????
#work to find best PCA value before knn and k loop 
```{r}
expr.dat = readRDS("/Users/hannagrossman/Desktop/BDSI Research/expression.rds")
screen.dat = readRDS("/Users/hannagrossman/Desktop/BDSI Research/screening.rds")
classes.1014 <- subset(screen.dat, DRUG_ID_lib == "1014")[,c("CL","EFFECT")]
expr.1014 <- expr.dat[as.character(classes.1014$CL),]
exprnew <- cbind(classes.1014, expr.1014)

x=exprnew[,3:17737]
y=exprnew[,2]
  
M <- 1:5
n <- nrow(x)
Q <- 1:5

#cv.error <- vector(mode = "numeric", length=10)

#best.k <- vector(mode="numeric", length=10)

#for(i in 1:10){
  
  train.ix <- sample(1:n, (19/20)*n)
  xtrain = x[train.ix,]
  ytrain = y[train.ix]
  xtest = x[-train.ix,]
  ytest = y[-train.ix]
  
  m.error <- matrix(, nrow=5, ncol=5) 
  
  for(m in M){
    
    #inside.error <- vector(mode="numeric", length= 10)
    
     for (q in Q){ 
       inside.error <- vector(mode="numeric", length= 10)
        for(j in 1:10){
          n2 <- nrow(xtrain)
          train2.ix <- sample(1:n2, (19/20)*n2)
          xtrain2 = xtrain[train2.ix,]
          ytrain2 = ytrain[train2.ix]
          xvalidate = xtrain[-train2.ix,]
          yvalidate = ytrain[-train2.ix]
      
      #pca on train 2; select q top components and create new data.frame 
      #then also create a new test data.frame w same linear combos as train one
      #use these new test and train data.frames to find best k below 
      #center data
          center.xtrain2 <- scale(xtrain2)
          center.xvalidate <- scale(xvalidate)
      #do prcomp to get rotation matrix on train dataset
          pca <- prcomp(center.xtrain2, center = TRUE)
          rotation <- as.data.frame(pca$rotation)[,c(1:q)]
          pca_xtrain2 <- as.matrix(center.xtrain2) %*% as.matrix(rotation)
          pca_xvalidate <- as.matrix(center.xvalidate) %*% as.matrix(rotation)
    
          fit <- knn(train = pca_xtrain2, test = pca_xvalidate, cl=ytrain2, k=m)
          inside.error[j]<- mean(fit != yvalidate)
        }
       m.error[m,q]=mean(inside.error)
       #m.error <- append(m.error, mean(inside.error))
     }
  }
#   
#   idx <- which.min(m.error)
#   best.k[i] <- M[idx]
#   
#   fit.train <- knn(train = xtrain, test = xtest, cl = ytrain, k = M[idx])
#   cv.error[i] <- mean(fit.train != ytest)
# }
# 
# mean(cv.error)
# best.k
#       
    

```

#for loop within a for loop for knn - choosing k and then performing knn - need to first make p smaller before performing knn 
```{r}
library(class)

x=pc.scores[,3:12]
y=pc.scores[,2]
  
M <- 1:100
n <- nrow(x)

cv.error <- vector(mode = "numeric", length=10)

best.k <- vector(mode="numeric", length=10)

for(i in 1:10){
  
  train.ix <- sample(1:n, (19/20)*n)
  xtrain = x[train.ix,]
  ytrain = y[train.ix]
  xtest = x[-train.ix,]
  ytest = y[-train.ix]
  
  m.error <- vector(mode = "numeric")
  
  for(m in M){
    
    inside.error <- vector(mode="numeric", length= 10)
    
    for(j in 1:10){
      n2 <- nrow(xtrain)
      train2.ix <- sample(1:n2, (19/20)*n2)
      xtrain2 = xtrain[train2.ix,]
      ytrain2 = ytrain[train2.ix]
      xvalidate = xtrain[-train2.ix,]
      yvalidate = ytrain[-train2.ix]
      
      fit <- knn(train = xtrain2, test = xvalidate, cl=ytrain2, k=m)
      inside.error[j]<- mean(fit != yvalidate)
    }
    
    m.error <- append(m.error, mean(inside.error))
  }
  
  idx <- which.min(m.error)
  best.k[i] <- M[idx]
  
  fit.train <- knn(train = xtrain, test = xtest, cl = ytrain, k = M[idx])
  cv.error[i] <- mean(fit.train != ytest)
}

mean(cv.error)
best.k

#1-cv.error
  

#for PCA = 10:
  #0.1461538
  #8 27 42 23 16  5 11 15  6 26

#for PCA =25:
  #0.1538462
  #8  1  1  3  8  3  6 10  3  5

#for PCA = 50: 
  #0.1923077
  #3 3 3 2 1 2 1 3 1 3

#for PCA = 100:
  #0.3538462
  #3  1  2 11  3  2  2  2  1  1
```

#redo to get best k value and then find best knn
```{r}
#inner part of loop above? 


```

#PCA initial analysis  
```{r}
#read in screen and expression data

#subset to drug 1014
classes.1014 <- subset(screen.dat, DRUG_ID_lib == "1014")[,c("CL", "EFFECT")]
expr.1014 <- expr.dat[as.character(classes.1014$CL),]
dim(expr.1014)

#center the expression data by column means
center.expr <- scale(expr.1014)

#now do PCA - here we choose 3 components but we probably wann test 1-100 to see what is best - do loop 10 times (10 diff train and test sets)
pca <- svd(center.expr, nu=3, nv = 0)

pc.scores <- as.data.frame(diag(pca$d) %*% pca$u)
pc.scores <- cbind(classes.1014, pc.scores)
dim(pc.scores)
head(pc.scores)
```

#plotting PCA 
```{r}
library(ggplot2)
ggplot(data=pc.scores)+ geom_point(aes(x=V1, y=V2, color = EFFECT)) + labs(x="PC1", y="PC2")

#can do this for v1&v3, v2&v3 
#here we see that a large source of the variation in the data is related to the effectiveness of drug 1014 (bc we see a separation between the two groups )

```

#PCR
```{r}
#find the first k principal components of the features & use those principal components as predictors and regress the response onto them 

library("pls")

#split data into train and test sets 
n <- dim(center.expr)[1]
train.index <- sample(1:n, n/2)

#training set
x.train <- center.expr[train.index,]
y.train <- classes.1014[train.index,]
train.dat <- data.frame(class = y.train$EFFECT, x.train)

#test set
x.test <- center.expr[-train.index,]
y.test <- classes.1014[-train.index,]
test.dat <- data.frame(class = y.test$EFFECT, x.test)
dim(x.train)
dim(y.train)

#use a subset of the data for this example (is this necessary? does the pcr test not do this already?)
train.dat<- train.dat[,1:10000]
test.dat <- test.dat[,1:10000]
train.dat[1:6,1:6]

#do PCR for the training data
#train.dat <- as.data.frame(train.dat)
pcr.fit <- pcr(class~., data=train.dat, scale = TRUE, ncomp =10)

#now use that fit to get predictions for the testing data
pcr.pred <- predict(pcr.fit, test.dat, ncomp = 10)
preds <- ifelse(pcr.pred >= 0.5, 1, 0)
mean(preds == y.test$EFFECT)
head(preds)

```

#PCR Loop attempt 
```{r}
expr.dat[1:6,1:6]
expr.dat = readRDS("/Users/hannagrossman/Desktop/BDSI Research/expression.rds")
screen.dat = readRDS("/Users/hannagrossman/Desktop/BDSI Research/screening.rds")

classes.1014 <- subset(screen.dat, DRUG_ID_lib == "1014")[,c("CL", "EFFECT")]
expr.1014 <- expr.dat[as.character(classes.1014$CL),]
center.expr <- scale(expr.1014)

  #should be 1:100 below
M <- 1:10
n <- nrow(center.expr)

cv.error <- vector(mode = "numeric", length=10)

best.k <- vector(mode="numeric", length=10)

for(i in 1:10){
  
  train.index <- sample(1:n, (19/20)*n)  #better to do 1/2 or 19/20? 
  x.train <- center.expr[train.index,]
  y.train <- classes.1014[train.index,]
  train.dat <- data.frame(class = y.train$EFFECT, x.train) #data.frame
  x.test <- center.expr[-train.index,]
  y.test <- classes.1014[-train.index,]
  test.dat <- data.frame(class = y.test$EFFECT, x.test) #data.frame
  
  m.error <- vector(mode = "numeric")
  
  for(m in M){
    inside.error <- vector(mode="numeric", length= 10)
  #  
    for(j in 1:10){
      #taking a sample of our train data (splitting past train data into train and test data sets)
      n2 <- nrow(x.train)
      train2.ix <- sample(1:n2, (19/20)*n2)
      xtrain2 = x.train[train2.ix,]
      ytrain2 = y.train[train2.ix,]
      train2.dat <- data.frame(class2 = ytrain2$EFFECT, xtrain2) #data.frame
      xvalidate = x.train[-train2.ix,]
      yvalidate = y.train[-train2.ix,]
      validate.dat <- data.frame(class2 = yvalidate$EFFECT, xvalidate) #data.frame
      
      pcr2.fit <- pcr(class2~., data=train2.dat, scale = TRUE, ncomp =m)
      pcr2.pred <- predict(pcr2.fit, validate.dat, ncomp = m)
      preds2 <- ifelse(pcr2.pred >= 0.5, 1, 0)
      inside.error[j] <- mean(preds2 == yvalidate$EFFECT)
    }
    
    m.error <- append(m.error, mean(inside.error))
  }
  
  idx <- which.min(m.error)
  best.k[i] <- M[idx]
  
  pcr.train <- pcr(class~., data=train.dat, scale = TRUE, ncomp= M[idx])
  pcr.pred <- predict(pcr.train, test.dat, ncomp = M[idx])
  preds <- ifelse(pcr.pred >= 0.5, 1, 0)
  cv.error[i] <- mean(preds == y.test$EFFECT)
}

# #############test above pcr
# pcr.fit <- pcr(class~., data=train.dat, scale = TRUE, ncomp =10)
# 
# #now use that fit to get predictions for the testing data
# pcr.pred <- predict(pcr.fit, test.dat, ncomp = 10)
# preds <- ifelse(pcr.pred >= 0.5, 1, 0)
# mean(preds == y.test$EFFECT)
# head(preds)

```

